Anony\emph{mice}d shareable data: Using \emph{mice} to create and analyze multiply imputed synthetic data sets

### 2021-11-01 manuscript/Manuscript.tex
Famous final words

% Authors, for the paper (add full first names)
\Author{Thom Volker$^{1,*,
\dagger}$\href{https://orcid.org/0000-0002-2408-7820}{\orcidicon}, Gerko
Vink$^{1,
\dagger}$\href{https://orcid.org/0000-0001-9767-1924}{\orcidicon}}

% Authors, for metadata in PDF
\AuthorNames{Thom Volker, Gerko Vink}

@@

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{Synthetic data sets can greatly improve the dissemination of
data and further analysis of private data. Generating and analyzing
synthetic data sets is straightforward, yet a synthetic data analysis
pipeline is seldom adopted by applied researchers. We outline a simple
procedure for generating and analyzing synthetic data sets with
\texttt{mice} in \texttt{R}. We demonstrate using simulations that the
analysis results obtained on synthetic data yields unbiased and valid
inferences, and leads to synthetic records that cannot be distinguished
from the true data records. The ease of use when synthesizing data with
\texttt{mice}, together with the validity of inferences obtained through
this procedure opens up a wealth of possibilities for data dissemination
and further research of initially private data.}

% Keywords
\keyword{mice; imputation; synthetic data.}

@@

Conceptually, the synthetic data framework is based upon the building
blocks of multiple imputation of missing data, as proposed by
\citet{rubin_multiple_1987}. Instead of replacing just the missing
values, one could easily replace any observed sensitive values by
multiple draws from the posterior predictive distribution - or a
non-Bayesian predictive distribution. We will refer to both of these as
predictive distributions. Similarly to multiple imputation of missing
data, the multiple synthetic data sets allow for correct statistical
inferences, despite the fact that the analyses do not use the ``true''
value. The analyses over multiple synthetic data sets should be pooled
into a single inference, so that the researcher can draw valid
conclusions from the pooled results. To that respect, the variance
should reflect the added variability that is induced by the imputation
procedure.

@@

The \texttt{mice} package \citep{mice} in \texttt{R} \citep{Rproject}
has been developed for multiple imputation to overcome problems related
to nonresponse. In that context, the aim is to replace missing values
with plausible values from the predictive distribution of that variable.
In \texttt{mice}, this is accomplished using fully conditional
specification (FCS) \citep{vanbuuren_fully_2006}, which breaks down the
multivariate distribution of the data
\(\textbf{Y} = (\textbf{Y}_{obs}, \textbf{Y}_{mis})\) into
\(j = 1, 2, \dots, k\) univariate conditional densities, where \(k\)
denotes the number of columns in the data. Using FCS, a model is
constructed for every incomplete variable and the missing values
\(Y_{j, mis}\) are then imputed with draws from the predictive
distribution of \(P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)\) on a
variable-by-variable basis. Note that the predictor matrix \(Y_{-j}\)
may contain yet imputed values from an earlier imputation step, and thus
will be updated after every iteration. This procedure is applied \(m\)
times, resulting in \(m\) completed data sets
\(\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})\),
with \(\textbf{D}^{(l)} = (\textbf{Y}_{obs}, Y^{(l)}_{mis})\).

@@

The imputation approach in \texttt{mice} can be easily extended to
generate synthetic values. Rather than imputing missing data, observed
values are then replaced by synthetic draws from the predictive
distribution. For simplicity, assume that the data is completely
observed (i.e., \(\textbf{Y} = \textbf{Y}_{obs}\)). Following the
notation of \citet{reiter_raghunathan_multiple_2007}, let, given \(n\)
observations, \(Z_i = 1\) if any of the values of unit
\(i = 1, 2, \dots, n\), are to be replaced by imputations, and
\(Z_i = 0\) otherwise, with \(\textbf{Z} = (Z_1, Z_2, \dots, Z_n)\).
Accordingly, the data consists of values that are to be replaced and
values that are to be kept (i.e.,
\(\textbf{Y} = (\textbf{Y}_{rep}, \textbf{Y}_{nrep})\). Now, instead of
imputing \(\textbf{Y}_{mis}\) with draws from the predictive
distribution of \(P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)\) as in the
missing data case, \(\textbf{Y}_{rep}\) is imputed from the posterior
distribution of
\(P(Y^{(l)}_{j, rep} | \textbf{Y}^{(l)}_{-j}, \textbf{Z}, \theta)\),
where \(l\) is an indicator for the synthetic data set
(\(l = 1, 2, \dots, m\)). The synthetic values that are imputed at an
earlier step can be used for imputing variable \(j\). This process
results in the synthetic data
\(\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})\).
Note that, given a completely observed data set, iterating over the
predictive distribution is not required.

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(mice)}
\NormalTok{syn }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(data.to.synthesize, }
            \AttributeTok{m =} \DecValTok{10}\NormalTok{,}
            \AttributeTok{maxit =} \DecValTok{1}\NormalTok{,}
            \AttributeTok{where =} \FunctionTok{matrix}\NormalTok{(}\ConstantTok{TRUE}\NormalTok{,}
                           \AttributeTok{nrow =} \FunctionTok{nrow}\NormalTok{(data.to.synthesize),}
                           \AttributeTok{ncol =} \FunctionTok{ncol}\NormalTok{(data.to.synthetize)))}
\end{Highlighting}
\end{Shaded}

Here, the argument \texttt{where} requires a matrix of the same
dimensions as the data, (i.e., a \(n \times k\) matrix) containing
logicals \(z_{ij}\) that indicate which cells are selected to have their
values replaced by draws from the predictive distribution. This approach
allows to replace a subset of the observed data (e.g., by only
specifying those cells that are to be replaced as \texttt{TRUE} in the
\texttt{where}-matrix, leaving the rest to \texttt{FALSE}), or as in the
aforementioned example, the observed data as a whole, resulting in a
data set that partially or completely consists of synthetic data values.
Note that because the data is completely observed, iterating over the
predictive distribution is not required.

@@

Choosing an adequate imputation model to impute the data is paramount,
as a flawed imputation model may drastically impact the validity of
inferences \citep{grund2021using, jiang2021balancing}. Imputation models
should be as flexible as possible to capture most of the patterns in the
data, and to model possibly unanticipated data characteristics
\citep{murray_multiple_2018, rubin_18years_1996}. Generally, multiple
methods can be feasible, including joint modeling, sequential regression
and FCS, all with different parameter specifications, depending on the
imputation problem at hand
\citep{ludtke2020regression, yucel2018sequential, murray2016multiple}.
Various simulations have shown that the most accurate imputation
procedure may vary over different imputation problems
\citep{hanne_review_2021}. Hence, a one-size-fits-all solution does
probably not exist.

However, one of the methods that has shown a reasonable and consistent
performance over multiple simulations using FCS in combination with
classification and regression trees
\citep[CART;][]{breiman_cart_1984, reiter_cart_2005, burgette_reiter_cart_2010, doove_buuren_recursive_2014, raab_practical_2016}.
Due to its non-parametric nature, CART allows to model more complex
patterns (e.g., interactions and non-linear relationships) in the data.
Loosely speaking, CART sequentially splits the predictor space into
non-overlapping regions in such a way that the within-region variance is
as small as possible after every split. As such, CART does not impose
any parametric distribution on the data, making it a widely applicable
method that allows for a large variety of relationships within the data
\citep{islr_2013}. Given these appealing characteristics and the call
for the use of flexible methods when multiply imputing data, we will
focus our illustrations and evaluations of \texttt{mice} to method
\texttt{mice.impute.cart()}, realized by:

@@

In a nutshell, the above code shows the simplicity of creating
\(m = 10\) synthetic data sets using \texttt{mice}. In practice,
however, one should take some additional complicating factors into
account. For example, one should account for deterministic relations in
the data. Additionally, relations between variables may be described
best using a different model than \texttt{CART}. Such factors are data
dependent and should be considered by the imputer. In the next section,
we will describe how the \texttt{mice::boys} data
\citep{fredriks_boys_2000} can be adequately synthesized. Additionally,
we will show through simulations that this approach yields valid
inferences.

@@

Unfortunately, this data set does not differ from the vast majority of
collected data sets, in the sense that it suffers from missing data. For
simplicity, we complete the missing values using the default
\texttt{mice} imputation model for all predictors except \texttt{bmi},
which is passively imputed using its deterministic relation with
\texttt{wgt} and \texttt{hgt}. Specifically, the imputed values are used
to calculate the exact \texttt{bmi} values that correspond with
\texttt{hgt} and \texttt{wgt}.

@@

Usually, one would draw samples from a population that can be
synthesized to evaluate the performance of the synthetisation methods.
As we only have access to a single sample, 1000 bootstrap samples of the
\texttt{boys} data have been synthesized with \(m = 5\) imputations for
every data cell to induce an appropriate amount of sampling variance.
This approach precludes us from knowing the \emph{true} data-generating
model, and in this sense provides a more stringent test of the
applicability of \texttt{mice} for real-life synthetisation procedures.
Specifically, every bootstrapped sample is treated as an actual sample
from a population that is synthesized using \texttt{mice}, using a model
that is built to approximate the true data-generating mechanisms as
closely as possible. As relationships in realistic data are generally
more complex than relationships in parametrically simulated data,
achieving good performance on a real data set is likely to be more
indicative of practical applicability than good performance on a
parametrically simulated data set.

@@

\[
T = \bar{u}_m + \frac{b_m}{m},
\]

with between-imputation variance

@@

\[
\bar{u}_m = \sum_{l = 1} \frac{u^{(l)}}{m},
\] where \(u^{(l)}\) denotes the variance estimate in the \(l\)th
synthetic data set. These pooling rules are implemented in \texttt{mice}
as the function \texttt{pool.syn}. \citet{reiter_kinney_2012} proved
that synthesizing data from the posterior predictive distribution is not
required for these pooling rules to hold. In fact, the parameters of an
imputation model can be fixed to their maximum likelihood estimates or
their posterior modes, and synthetic values can be drawn after plugging
these values into the imputation models. This observation as been taken
a step further by \citet{raab_practical_2016}, who developed pooling
rules that do not require multiple synthetic data sets and performed
highly similar to the aforementioned pooling rules. However, we adhere
to the originally proposed pooling rules by
\citet{reiter_partially_inference_2003}, who defined the total variance
in terms of sampling variance and between-imputation variance.

@@

We evaluate the synthetic data with respect to the \emph{true} data set
on the basis of three aims. We believe that every reliable and valid
data synthetisation effort in statistical data analysis should be able
to yield 1) unbiased univariate statistical properties, 2) unbiased
bivariate properties, 3) unbiased and valid multivariate inferences and
4) synthetic data that cannot be distinguished from real data. We
consider the evaluation of the synthetic data simulations in the above
order.

@@

The sampling distribution of the means of the continuous variables for
both the bootstrapped and the synthetic data can be found in Figure 1,
while the number of observations per category in the bootstrapped data
and the averaged number of observations per category over the synthetic
data sets can be found in Figure 2. Both figures show that the marginal
distributions of the bootstrapped data and the synthetic data are highly
similar. Additionally, the univariate descriptives for the original data
and the synthetic data can be found in Table 2.

@@

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/univariate-statistics-II-1.pdf}
\caption{Empirical sampling distribution of the number within each
category of the categorical variables in the bootstrapped and synthetic
boys data.}
\end{figure}

\begin{table}[ht]
\caption{Univariate descriptives for the true data and $m=5$ pooled univariate descriptives for the synthetic data over 1000 simulations. Variable names followed by a $^*$ are categorical.}
\centering
\begin{tabular}{rrrrrrrrr}
  \hline
 & n & mean & sd & median & min & max & skew & kurtosis \\ 
  \hline
original age & 748 & 9.16 & 6.89 & 10.50 & 0.04 & 21.18 & -0.03 & -1.56 \\ 
  synthetic age & 748 & 9.15 & 6.89 & 10.49 & 0.04 & 20.96 & -0.03 & -1.55 \\ 
  original hgt & 748 & 131.10 & 46.52 & 145.75 & 50.00 & 198.00 & -0.30 & -1.47 \\ 
  synthetic hgt & 748 & 131.06 & 46.50 & 145.32 & 50.69 & 197.16 & -0.30 & -1.47 \\ 
  original wgt & 748 & 37.12 & 26.03 & 34.55 & 3.14 & 117.40 & 0.38 & -1.03 \\ 
  synthetic wgt & 748 & 37.09 & 26.00 & 34.44 & 3.35 & 112.26 & 0.38 & -1.03 \\ 
  original bmi & 748 & 18.04 & 3.04 & 17.45 & 11.73 & 31.74 & 1.14 & 1.79 \\ 
  synthetic bmi & 748 & 18.05 & 3.08 & 17.48 & 11.49 & 32.37 & 1.11 & 1.85 \\ 
  original hc & 748 & 51.62 & 5.86 & 53.10 & 33.70 & 65.00 & -0.91 & 0.12 \\ 
  synthetic hc & 748 & 51.61 & 5.86 & 53.18 & 34.38 & 62.85 & -0.91 & 0.12 \\ 
  original gen* & 748 & 2.53 & 1.59 & 2.00 & 1.00 & 5.00 & 0.52 & -1.36 \\ 
  synthetic gen* & 748 & 2.53 & 1.59 & 2.00 & 1.00 & 5.00 & 0.52 & -1.35 \\ 
  original phb* & 748 & 2.75 & 1.86 & 2.00 & 1.00 & 6.00 & 0.56 & -1.25 \\ 
  synthetic phb* & 748 & 2.75 & 1.86 & 2.00 & 1.00 & 6.00 & 0.56 & -1.24 \\ 
  original tv & 748 & 8.43 & 8.12 & 3.00 & 1.00 & 25.00 & 0.85 & -0.78 \\ 
  synthetic tv & 748 & 8.42 & 8.11 & 3.19 & 1.00 & 25.00 & 0.85 & -0.77 \\ 
  original reg* & 748 & 3.02 & 1.14 & 3.00 & 1.00 & 5.00 & -0.08 & -0.77 \\ 
  synthetic reg* & 748 & 3.02 & 1.14 & 3.00 & 1.00 & 5.00 & -0.08 & -0.76 \\ 
   \hline
\end{tabular}
\end{table}

We see from Table 1 that the synthetic data estimates closely resemble
the true data estimates. All sample statistics of interest show
negligible bias over the 1000 synthetic data sets. Hence, univariately
the synthetisation model proves adequate.

@@

An often used bivariate statistic is Pearson's correlation coefficient.
When evaluating this correlation coefficient on the numeric columns in
the \texttt{boys} data set, we find that biases are very small. The
results are displayed in Table 3.

@@

We see that the correlations obtained from synthetic data are unbiased
with respect to the true data set. The largest absolute bias over 1000
simulations equals 0.013, indicating that the imputation model is
capable of preserving the bivariate relations in the data.

@@

However, when creating synthetic data with \texttt{mice}, close
attention should be paid to three distinct aspects of the data:
incorporating the additional uncertainty that is due to synthesizing
(part of) the data, potential disclosure risks that remain after
synthesizing the data, and the utility of the synthetic data. First, the
procedure of generating multiple synthetic sets may seem overly
complicated. We would like to emphasize that analyzing a single
synthesized set, while perhaps unbiased, would underestimate the
variance properties that are so important in drawing statistical
inferences from finite data sets. After all, we are often not interested
in the sample at hand, but aim to make inferences about the underlying
data generating mechanism as reflected in the population. Properly
capturing the uncertainty of synthetic data sets, just like with
incomplete data sets, is therefore paramount. To achieve this, we
adopted a bootstrap scheme in our simulations, to represent sampling
from a superpopulation. However, when a single sample should be the
reference, in the sense that the sample reflects the population one
wishes to make inferences about, adjusted pooling rules are required
that are similar to the procedure outlined in \citet{vink2014pooling}.
The corresponding pooling rules have not been derived yet and the
incorporation in data analysis workflows would require proper attention
from developers alike.

@@

While \texttt{synthpop} is capable to synthesize regardless of missing
data, it is not developed to solve missingness problems. If the data for
synthetisation contains missing values, the missingness will be
considered as a property of the data, and the corresponding synthetic
data sets will contain missing values as well. If the goal is to solve
for the missingness and synthesize the data set, a two-step approach
must be used, in which the missing values are imputed using a different
package, like \texttt{mice}. After this, \texttt{synthpop} can be used
to synthesize the imputed data. Then, given \(m\) multiple imputations
and \(r\) synthetisations, at least \(m \times r\) synthetic data sets
are in order, after which the data can be analysed and the analyses can
be pooled using rules developed by \citet{reiter2004simultaneous}.
However, such an approach would be computationally inefficient and
practically cumbersome, because the architectures of both packages
differ.

The appealing nature of \texttt{mice} therefore lays in its ability to
solve both the missingness problem and create synthetic versions of the
data. Eventually, the flexibility with \texttt{mice} is that both
unobserved and observed data values could be synthesized at once,
without the need for a two-step approach. Using \texttt{mice}, \(m\)
synthetic sets would be sufficient. However, as of today, no pooling
rules for one-step imputation of missingness and synthetisation have
been developed. The derivation of those would yield a massive
theoretical improvement to further reduce the burden of creating
synthetic data sets.

>>>>

% Authors, for the paper (add full first names)
\Author{Thom Benjamin Volker$^{1,*,
\dagger}$\href{https://orcid.org/0000-0002-2408-7820}{\orcidicon}, Gerko
Vink$^{1,
\dagger}$\href{https://orcid.org/0000-0001-9767-1924}{\orcidicon}}

% Authors, for metadata in PDF
\AuthorNames{Thom Benjamin Volker, Gerko Vink}

@@

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{Synthetic data sets simultaneously allow for the dissemination
of research data while protection the privacy and confidentiality of
respondents. Generating and analyzing synthetic data sets is
straightforward, yet a synthetic data analysis pipeline is seldom
adopted by applied researchers. We outline a simple procedure for
generating and analyzing synthetic data sets with the multiple
imputation software \texttt{mice} in \texttt{R}. We demonstrate using
simulations that the analysis results obtained on synthetic data yield
unbiased and valid inferences, and leads to synthetic records that
cannot be distinguished from the true data records. The ease of use when
synthesizing data with \texttt{mice}, together with the validity of
inferences obtained through this procedure opens up a wealth of
possibilities for data dissemination and further research of initially
private data.}

% Keywords
\keyword{mice; multiple imputation; synthetic data; statistical
disclosure control; privacy.}

@@

Conceptually, the synthetic data framework is based upon the building
blocks of multiple imputation of missing data, as proposed by
\citet{rubin_multiple_1987}. Instead of replacing just the missing
values, one could easily replace any observed sensitive values by
multiple draws from the posterior predictive distribution - or a
non-Bayesian predictive distribution (both of these will be referred to
as predictive distributions). Similarly to multiple imputation of
missing data, the multiple synthetic data sets allow for correct
statistical inferences, despite the fact that the analyses do not use
the \emph{true} value. The analyses over multiple synthetic data sets
should be pooled into a single inference, so that the researcher can
draw valid conclusions from the pooled results. To that respect, the
variance should reflect the added variability that is induced by the
imputation procedure.

@@

Fully conditional specification (FCS) has been implemented in the
\texttt{mice} package \citep{mice, vanbuuren_fully_2006} in \texttt{R}
\citep{Rproject}, which has been developed for multiple imputation to
overcome problems related to nonresponse. In that context, the aim is to
replace missing values with plausible values from the predictive
distribution of that variable. This is achieved by breaking down the
multivariate distribution of the data
\(\textbf{Y} = (\textbf{Y}_{obs}, \textbf{Y}_{mis})\) into
\(j = 1, 2, \dots, k\) univariate conditional densities, where \(k\)
denotes the number of columns in the data. Using FCS, a model is
constructed for every incomplete variable and the missing values
\(Y_{j, mis}\) are then imputed with draws from the predictive
distribution of \(P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)\) on a
variable-by-variable basis. Note that the predictor matrix \(Y_{-j}\)
may contain yet imputed values from an earlier imputation step, and thus
will be updated after every iteration. This procedure is applied \(m\)
times, resulting in \(m\) completed data sets
\(\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})\),
with \(\textbf{D}^{(l)} = (\textbf{Y}_{obs}, \textbf{Y}^{(l)}_{mis})\).

@@

The imputation approach in \texttt{mice} can be easily extended to
generate synthetic values. Rather than imputing missing data, observed
values are then replaced by synthetic draws from the predictive
distribution. For simplicity, assume that the data is completely
observed (i.e., \(\textbf{Y} = \textbf{Y}_{obs}\)). Following the
notation of \citet{reiter_raghunathan_multiple_2007}, let, given \(n\)
observations, \(Z_i = 1\) if any of the values of unit
\(i = 1, 2, \dots, n\), are to be replaced by imputations, and
\(Z_i = 0\) otherwise, with \(Z = (Z_1, Z_2, \dots, Z_n)\). Accordingly,
the data consists of values that are to be replaced and values that are
to be kept (i.e.,
\(\textbf{Y} = (\textbf{Y}_{rep}, \textbf{Y}_{nrep})\). Now, instead of
imputing \(\textbf{Y}_{mis}\) with draws from the predictive
distribution of \(P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)\) as in the
missing data case, \(\textbf{Y}_{rep}\) is imputed from the posterior
distribution of \(P(Y^{(l)}_{j, rep} | \textbf{Y}_{-j}, Z, \theta)\),
where \(l\) is an indicator for the synthetic data set
(\(l = 1, 2, \dots, m\)). This process results in the synthetic data
\(\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})\).

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{syn }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(data.to.synthesize, }
            \AttributeTok{m =} \DecValTok{10}\NormalTok{,}
            \AttributeTok{maxit =} \DecValTok{1}\NormalTok{,}
            \AttributeTok{where =} \FunctionTok{matrix}\NormalTok{(}\ConstantTok{TRUE}\NormalTok{,}
                           \AttributeTok{nrow =} \FunctionTok{nrow}\NormalTok{(data.to.synthesize),}
                           \AttributeTok{ncol =} \FunctionTok{ncol}\NormalTok{(data.to.synthetize)))}
\end{Highlighting}
\end{Shaded}

Here, the argument \texttt{where} requires a matrix of the same
dimensions as the data, (i.e., a \(n \times k\) matrix) containing
logicals \(z_{ij}\) that indicate which cells are selected to have their
values replaced by draws from the predictive distribution. This approach
allows to replace a subset of the observed data (e.g., by specifying
only those cells that are to be replaced as \texttt{TRUE} in the
\texttt{where}-matrix, leaving the rest to \texttt{FALSE}), or as in the
aforementioned example, the observed data as a whole, resulting in a
data set that partially or completely consists of synthetic data values.
Note that because the data is completely observed, iterating over the
predictive distribution is not required.

@@

Choosing an adequate imputation model to impute the data is paramount,
as a flawed imputation model may drastically impact the validity of
inferences \citep{grund2021using, jiang2021balancing}. Imputation models
should be as flexible as possible to capture most of the patterns in the
data, and to model possibly unanticipated data characteristics
\citep{murray_multiple_2018, rubin_18years_1996}. Generally, multiple
methods can be feasible, including joint modeling, sequential regression
and FCS, all with different parameter specifications, depending on the
imputation problem at hand
\citep{ludtke2020regression, yucel2018sequential, murray2016multiple}.
Various simulations have shown that the most accurate imputation
procedure may vary over different imputation problems \citep[e.g., see][
for an overview]{hanne_review_2021}. Hence, a one-size-fits-all solution
does probably not exist.

However, one of the methods that has shown a reasonable and consistent
performance over multiple simulations
\citep[e.g.,][]{reiter_cart_2005, burgette_reiter_cart_2010, doove_buuren_recursive_2014, raab_practical_2016}
is FCS in combination with classification and regression trees
\citep[CART;][]{breiman_cart_1984}. Due to its non-parametric nature,
CART allows to model complex patterns (e.g., interactions and non-linear
relationships) in the data. Loosely speaking, CART sequentially splits
the predictor space into non-overlapping regions in such a way that the
within-region variance is as small as possible after every split. As
such, CART does not impose any parametric distribution on the data,
making it a widely applicable method that allows for a large variety of
relationships within the data to be modeled \citep{islr_2013}. Given
these appealing characteristics and the call for the use of flexible
methods when multiply imputing data, we will focus our illustrations and
evaluations of \texttt{mice} to method \texttt{mice.impute.cart()},
realized by:

@@

In a nutshell, the above code shows the simplicity of creating
\(m = 10\) synthetic data sets using \texttt{mice}. In practice,
however, one should take some additional complicating factors into
account. For example, one should account for deterministic relations in
the data. Additionally, relations between variables may be described
best using a different model than \texttt{CART}. Such factors are data
dependent and should be considered by the imputer. In the next section,
we will describe how a completely observed version of the
\texttt{mice::boys} data \citep{fredriks_boys_2000} can be adequately
synthesized. Additionally, we will show through simulations that this
approach yields valid inferences.

@@

Unfortunately, this data set does not differ from the vast majority of
collected data sets, in the sense that it suffers from missing data. For
simplicity, we solve the missingness problem using a single imputation
with the default \texttt{mice} imputation model for all predictors
except \texttt{bmi}, which is passively imputed using its deterministic
relation with \texttt{wgt} and \texttt{hgt}. Specifically, the imputed
values are used to calculate the exact \texttt{bmi} values that
correspond with \texttt{hgt} and \texttt{wgt}.

@@

Usually, one would draw samples from a population that can be
synthesized to evaluate the performance of the synthetisation methods.
As we only have access to a single sample, 1000 bootstrap samples of the
\texttt{boys} data have been synthesized with \(m = 5\) imputations for
every data cell to induce an appropriate amount of sampling variance.
This approach precludes us from knowing the \emph{true} data-generating
model, and in this sense provides a more stringent test of the
applicability of \texttt{mice} for real-life synthetisation procedures.
Specifically, every bootstrapped sample is treated as an actual sample
from a population. These bootstrap samples will be synthesized using
\texttt{mice}, using a model that is built to approximate the true
data-generating mechanisms as closely as possible. As relationships in
realistic data are generally more complex than relationships in
parametrically simulated data, achieving good performance on a real data
set is likely to be more indicative of practical applicability than good
performance on samples from a known multivariate probability density
function.

@@

\[
T = \bar{u}_m + \frac{b_m}{m},
\] with between-imputation variance

@@

\[
\bar{u}_m = \sum_{l = 1}^m \frac{u^{(l)}}{m},
\] where \(u^{(l)}\) denotes the variance estimate in the \(l\)th
synthetic data set. These pooling rules are implemented in \texttt{mice}
as the function \texttt{pool.syn}. \citet{reiter_kinney_2012} proved
that synthesizing data from the posterior predictive distribution is not
required for these pooling rules to hold. In fact, the parameters of an
imputation model can be fixed to their maximum likelihood estimates or
their posterior modes, and synthetic values can be drawn after plugging
these values into the imputation models. This observation as been taken
a step further by \citet{raab_practical_2016}, who developed pooling
rules that do not require multiple synthetic data sets and performed
highly similar to the aforementioned pooling rules. However, because
these pooling rules can only be used when the complete data (rather than
a subset of the data) is synthesized, we adhere to the originally
proposed pooling rules by \citet{reiter_partially_inference_2003}, who
defined the total variance in terms of sampling variance and
between-imputation variance.

@@

The synthetic data is evaluated with respect to the \emph{true} data set
on the basis of three aims. We believe that every reliable and valid
data synthetisation effort in statistical data analysis should be able
to yield 1) unbiased univariate statistical properties, 2) unbiased
bivariate properties, 3) unbiased and valid multivariate inferences and
4) synthetic data that cannot be distinguished from real data. We
consider the evaluation of the synthetic data simulations in the above
order.

@@

The univariate descriptives for the original data and the synthetic data
can be found in Table 2. From this table, it can be seen that the
synthetic data estimates closely resemble the true data estimates. All
sample statistics of interest show negligible bias over the 1000
synthetic data sets.

The sampling distribution of the means of the continuous variables for
both the bootstrapped and the synthetic data can be found in Figure 1,
while the number of observations per category in the bootstrapped data
and the averaged number of observations per category over the synthetic
data sets can be found in Figure 2.

Both figures show that the marginal distributions of the bootstrapped
data and the synthetic data are highly similar. The sampling
distributions of these parameters hardly seem to differ. Hence, both the
numeric and graphical results show that univariately the synthetisation
model proves adequate.

@@

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/univariate-statistics-II-1.pdf}
\caption{Empirical sampling distribution of the number of observations
within each category of the categorical variables in the bootstrapped
and synthetic boys data.}
\end{figure}

\begin{table}[ht]
\caption{Univariate descriptives for the true data and $m=5$ pooled univariate descriptives for the synthetic data over 1000 simulations. Variable names followed by an asterix ($^*$) are categorical.}
\centering
\begin{tabular}{rrrrrrrrr}
  \hline
 & n & mean & sd & median & min & max & skew & kurtosis \\ 
  \hline
original age & 748 & 9.16 & 6.89 & 10.50 & 0.04 & 21.18 & -0.03 & -1.56 \\ 
  synthetic age & 748 & 9.15 & 6.89 & 10.49 & 0.04 & 20.96 & -0.03 & -1.55 \\ 
  original hgt & 748 & 131.10 & 46.52 & 145.75 & 50.00 & 198.00 & -0.30 & -1.47 \\ 
  synthetic hgt & 748 & 131.06 & 46.50 & 145.32 & 50.69 & 197.16 & -0.30 & -1.47 \\ 
  original wgt & 748 & 37.12 & 26.03 & 34.55 & 3.14 & 117.40 & 0.38 & -1.03 \\ 
  synthetic wgt & 748 & 37.09 & 26.00 & 34.44 & 3.35 & 112.26 & 0.38 & -1.03 \\ 
  original bmi & 748 & 18.04 & 3.04 & 17.45 & 11.73 & 31.74 & 1.14 & 1.79 \\ 
  synthetic bmi & 748 & 18.05 & 3.08 & 17.48 & 11.49 & 32.37 & 1.11 & 1.85 \\ 
  original hc & 748 & 51.62 & 5.86 & 53.10 & 33.70 & 65.00 & -0.91 & 0.12 \\ 
  synthetic hc & 748 & 51.61 & 5.86 & 53.18 & 34.38 & 62.85 & -0.91 & 0.12 \\ 
  original gen* & 748 & 2.53 & 1.59 & 2.00 & 1.00 & 5.00 & 0.52 & -1.36 \\ 
  synthetic gen* & 748 & 2.53 & 1.59 & 2.00 & 1.00 & 5.00 & 0.52 & -1.35 \\ 
  original phb* & 748 & 2.75 & 1.86 & 2.00 & 1.00 & 6.00 & 0.56 & -1.25 \\ 
  synthetic phb* & 748 & 2.75 & 1.86 & 2.00 & 1.00 & 6.00 & 0.56 & -1.24 \\ 
  original tv & 748 & 8.43 & 8.12 & 3.00 & 1.00 & 25.00 & 0.85 & -0.78 \\ 
  synthetic tv & 748 & 8.42 & 8.11 & 3.19 & 1.00 & 25.00 & 0.85 & -0.77 \\ 
  original reg* & 748 & 3.02 & 1.14 & 3.00 & 1.00 & 5.00 & -0.08 & -0.77 \\ 
  synthetic reg* & 748 & 3.02 & 1.14 & 3.00 & 1.00 & 5.00 & -0.08 & -0.76 \\ 
   \hline
\end{tabular}
\end{table}

@@

An often used bivariate statistic is Pearson's correlation coefficient.
When evaluating this correlation coefficient on the numeric columns in
the \texttt{boys} data set, the differences between the correlations in
the synthetic and bootstrapped data are very small. These results are
displayed in Table 3.

@@

The correlations obtained from synthetic data are unbiased with respect
to the true data set. The largest bias over 1000 simulations equals
0.013, indicating that the imputation model is capable of preserving the
bivariate relations in the data.

@@

When creating synthetic data with \texttt{mice}, close attention should
be paid to three distinct factors: (i) the additional uncertainty that
is due to synthesizing (part of) the data should be incorporated, (ii)
potential disclosure risks that remain after synthesizing the data
should be assessed, and (iii) the utility of the synthetic data should
be examined. First, the procedure of generating multiple synthetic sets
may seem overly complicated. We would like to emphasize that analyzing a
single synthesized set, while perhaps unbiased, would underestimate the
variance properties that are so important in drawing statistical
inferences from finite data sets. After all, we are often not interested
in the sample at hand, but aim to make inferences about the underlying
data generating mechanism as reflected in the population. Properly
capturing the uncertainty of synthetic data sets, just like with
incomplete data sets, is therefore paramount. To achieve this, we
adopted a bootstrap scheme in our simulations, to represent sampling
from a superpopulation. However, when a single sample should be the
reference, in the sense that the sample reflects the population one
wishes to make inferences about, adjusted pooling rules are required
that are similar to the procedure outlined in \citet{vink2014pooling}.
The corresponding pooling rules have not been derived yet and the
incorporation in data analysis workflows would require proper attention
from developers alike.

@@

While \texttt{synthpop} is capable to synthesize regardless of missing
data, it is not developed to solve missingness problems. If the data for
synthetisation contains missing values, the missingness will be
considered as a property of the data, and the corresponding synthetic
data sets will contain missing values as well. If the goal is to solve
for the missingness and synthesize the data set, a two-step approach
must be used, in which the missing values are imputed using a different
package, like \texttt{mice}. Subsequently, \texttt{synthpop} can be used
to synthesize the imputed data. Then, given \(m\) multiple imputations
and \(r\) synthetisations, at least \(m \times r\) synthetic data sets
are in order, after which the data can be analysed and the analyses can
be pooled using rules developed by \citet{reiter2004simultaneous}.
However, such an approach would be computationally inefficient and
practically cumbersome, because the architectures of both packages
differ.

The appealing nature of \texttt{mice} therefore lays in its ability to
both solve the missingness problem and create synthetic versions of the
data. Eventually, the flexibility with \texttt{mice} is that unobserved
and observed data values could be synthesized at once, without the need
for a two-step approach. Using \texttt{mice}, \(m\) synthetic sets would
be sufficient. However, as of today, no pooling rules for one-step
imputation of missingness and synthetisation have been developed. The
derivation of those would yield a massive theoretical improvement to
further reduce the burden of creating synthetic data sets.

That said, the fields of synthetic data and multiple imputation are very
much in development. Although efforts have been made to bridge gaps
between these fields, such as in the \texttt{miceadds} package
\citep{miceadds_3.11-6} which links functionality between between
\texttt{synthpop} and \texttt{mice}, further improvements could be made
in that respect. Specifically, \texttt{synthpop} contains a
comprehensive suite of functions that can be used to assess the quality
of synthetic data \citep[e.g., see][]{snoke2018general}, while
\texttt{mice} contains methodology to allow for flexible synthesis of
intricate data structures, such as multilevel data. Being able to use
such functionality interchangeably would greatly benefit data analysts.
